Dark matter halo shape at $z=0$ in the Auriga simulations: radial evolution and alignment with the stellar disk

### 2019-07-19 Document/paper.tex
referencias

This ellipsoidal shape is mostly due to the anisotropical and
clumpy accretion of matter influenced by environmental structures.
Numerical studies how that the shape has a strong mass dependence
\citep{Allgood_et_al._2006}, halos are also rounder at the outerskirts
than at the inner part. 
Shape also evolves with cosmic time, halos get
rounder as they evolve.

\citep{Dubinski91}

Observationally some studies prefer oblate (i.e. a=b>c) configurations at small
distances around $\leq 20$ kpc
\citep[see][]{Law_and_Majewski_2010,Bovy_et_el._2016,Loebman_et_al._2012,Olling_and_Merrifield_2000,Banerjee_and_Chanda_2011} 
and more triaxial and prolate configurations on the outter distances
$\geq 20$ kpc 
\citep[see][]{Vera-Ciro_and_Helmi_2013,Law_and_Majewski_2009,Deg_and_Widrow_2013,Banerjee_and_Chanda_2011}.
However, some  studies are inclined towards prolate configurations even at the inner
parts of the halo \citep[see][]{Bowden_et_al._2016}, and
although it previously seemed that a triaxial DM halo on the
outerskirts would be necessary to fully explain the characterization
of the Sagittarius stream \citep{Law_and_Majewski_2009}, recent studies
questioned this claim by reporting inconsistencies with narrow stellar
streams \citet{Pearson_et_al._2015} or finding that
the relaxation of other constraints may make this claim unnecessary
\citet{Ibata_et_al._2013}.

In simulations there is strong evidence claiming that the presence of
baryons produces axisymmetrical halos.  
For instance, some studies have shown that the DM halo shape must be
axisymmetrical to ensure the stability of a hydrodynamical disk
embeded in a static DM halo. 
Other have studied this rounding effect by simulating the disk as rigid
potential inside an N-body triaxial DM
halo \cite{Debattista_et_al._2008,Debattista_et_al._2013,Kazantzidis_et_al._2010}
finding that the halo responds to the disk by becoming less triaxial.

Rounding effect of baryons
\citep{Dubinski94}

The caveat of the studies mentioned above is that they do not
follow baryons in the whole cosmological context. 
Other studies overcome this limitation by using resimulations 
\citep{Abadi_et_al._2010,Bryan_et_al._2013} finding that the
feeback related to star formation in the disk drives the strenght of
the round effect.

Abadi2010 -> well aligned countours.

Recently \cite{2018arXiv180907255C} made a study in a cosmological
simulation to compare the effect of including baryons. They do find,
on average, rounder halo shapes once hydrodynamic effects are
included, but it is uncertain the strenght of this statistical effect
on galaxies similar to the MW.

\citep{Pedrosa10}

@@

Our second results deals with the disk-halo shape alignment as a function of
radius. 
In concordance with previous results in DM only simulations we find
that the shells of halo shape are well aligned at different radii.
However, in the presences of baryons these shells twist as a function
of radius.
At a radius of $0.25R_{200}\approx 56$ kpc the alignment between the
stellar disk and the dark matter halo is almost perfect and degrades
at other radii. 
This twisting effect, the change of disk-halo alignment as
a function of radius, is robust across our sample of 30
halos at all radii, its strenght strongly correlating with halo
triaxiality measured at $0.12R_{200}$.

\cite{Debattista08} reported a similar twisting effect, but only at radii smaller
than the disk radius. 
At radii larger than the disk radius, they find, the dark matter
shells stop twisting. 
\cite{JingSuto02} also reported shell alignment twisting in their dark
matter only simulations as measured in the direction changes of large
and medium shape axis. 
They only had three high resolution MW-like halos (with
$\sim 10^{6}$ inside the virial radius) and could not make a
statistical statement about the impact of this effect.
However, they the twisting as an artifact resulting from high values of the
$c/b$ ratios that make the determination of the medium axis boisy.
Could then high triaxility values explain the twisting as a result of
erratic axis directions? 
If the triaxiality were the main culprit behind the twisting, the
correlation between the two quantities should be constant at every
radii.
In Figure \ref{fig:alignment_correlations} we already showed that this
correlation strongly depends on radius, we conclude that high
triaxiality values cannot be the full explanation.
Furthermore the twisting happens in such a way that it allows an
almost perfect alignment between the stellar disk at $0.1R{200}$ and
the halo shape measured at $0.25R_{200}$

>>>>

These two questions have been difficult to address because for a long
time it was difficult to produce realistic galactic disks within the
CDM context.
The success of such enterprise has ben the result of high numerical
resolution and the understanding that baryonic effects such as stellar
feedback and black hole feedback play an important role in forming a
galactic disk resembling the Milky Way.

Along the way the experiments show that the baryonic effects also impact
the dark matter halo shape making the halo rounder than it otherwise
be in a dark matter only simulation
\citep{Dubinski94,Debattista08,Kazantzidis10,Abadi10,Bryan13,Chua19,Artale19}. 
High resolution simulations have also allowed detailed studies of the
alignment between the stellar disk and the dark matter halo
\cite{Bailin05,DeBuhr12,Debattista13,Gomez17}.

Observationally, some studies prefer oblate configurations at small
distances around $\leq 20$ kpc
\citep[see][]{Law_and_Majewski_2010,Bovy_et_el._2016,Loebman_et_al._2012,Olling_and_Merrifield_2000,Banerjee_and_Chanda_2011} 
and more triaxial and prolate configurations on the outter distances
$\geq 20$ kpc 
\citep[see][]{Vera-Ciro_and_Helmi_2013,Law_and_Majewski_2009,Deg_and_Widrow_2013,Banerjee_and_Chanda_2011}.

However, some  studies are inclined towards prolate configurations even at the inner
parts of the halo \citep[see][]{Bowden_et_al._2016}, and
although it previously seemed that a triaxial DM halo on the
outerskirts would be necessary to fully explain the characterization
of the Sagittarius stream \citep{Law_and_Majewski_2009}, recent studies
questioned this claim by reporting inconsistencies with narrow stellar
streams \citet{Pearson_et_al._2015} or finding that
the relaxation of other constraints may make this claim unnecessary
\citet{Ibata_et_al._2013}. 
axisymmetrical to ensure the stability of a hydrodynamical disk
embeded in a static DM halo.

@@

Our second results deals with the disk-halo shape alignment as a function of
radius. 
In concordance with previous results in DM only simulations we find
that the shells of halo shape are well aligned at different radii.
However, in the presences of baryons these shells twist as a function
of radius.
At a radius of $0.25R_{200}\approx 56$ kpc the alignment between the
stellar disk and the dark matter halo is almost perfect and degrades
at other radii. 
This twisting effect, the change of disk-halo alignment as
a function of radius, is robust across our sample of 30
halos at all radii.
However, its strenght is not the same for all galaxies and strongly
correlates with halo triaxiality measured at $0.12R_{200}$.

\cite{Bailin05} studied the disk halo alignmnent as a function of
radius.
\cite{Gomez17} studied this in Auriga.
\cite{Debattista13} reported a similar twisting effect, but only at radii smaller
than the disk radius. 
At radii larger than the disk radius, they find, the dark matter
shells stop twisting. 
\cite{JingSuto02} also reported shell alignment twisting in their dark
matter only simulations as measured in the direction changes of large
and medium shape axis. 
They only had three high resolution MW-like halos (with
$\sim 10^{6}$ inside the virial radius) and could not make a
statistical statement about the impact of this effect.
However, they the twisting as an artifact resulting from high values of the
$c/b$ ratios that make the determination of the medium axis boisy.
Could then high triaxility values explain the twisting as a result of
erratic axis directions? 
If the triaxiality were the main culprit behind the twisting, the
correlation between the two quantities should be constant at every
radii.
In Figure \ref{fig:alignment_correlations} we already showed that this
correlation strongly depends on radius, we conclude that high
triaxiality values cannot be the full explanation.
Furthermore the twisting happens in such a way that it allows an
almost perfect alignment between the stellar disk at $0.1R{200}$ and
the halo shape measured at $0.25R_{200}$
